---
title: "Assignment 3: Data Exploration"
author: "Xiangtian Wang"
output: pdf_document
geometry: margin=2.54cm
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Exploration. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.
8. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Salk_A03_DataExploration.Rmd") prior to submission.

The completed exercise is due on Tuesday, January 28 at 1:00 pm.


## Set up your R session

1. Check your working directory, load necessary packages (tidyverse), and upload two datasets: the ECOTOX neonicotinoid dataset (ECOTOX_Neonicotinoids_Insects_raw.csv) and the Niwot Ridge NEON dataset for litter and woody debris (NEON_NIWO_Litter_massdata_2018-08_raw.csv). Name these datasets "Neonics" and "Litter", respectively. 
```{r,  message=FALSE}
library (tidyverse)
Neoincs <- read.csv("../Data/Raw/ECOTOX_Neonicotinoids_Insects_raw.csv")
Litter <- read.csv ("../Data/Raw/NEON_NIWO_Litter_massdata_2018-08_raw.csv")

```

## Learn about your system

2. The neonicotinoid dataset was collected from the Environmental Protection Agency's ECOTOX Knowledgebase, a database for ecotoxicology research. Neonicotinoids are a class of insecticides used widely in agriculture. The dataset that has been pulled includes all studies published on insects. Why might we be interested in the ecotoxicologoy of neonicotinoids on insects? Feel free to do a brief internet search if you feel you need more background information. 

> Answer: Neonicotinoid has some by-product effect on bees, like honey-bee colony collapse disorder and causes a decrease in bird population due to the reduction of insects. However, the bee exposed to neonicotinoid, will not die immediately but has some damage to the neurons system. Scientists have a different conclusion about the ecotoxic of neonicotinoid. So we need to research more about it.  

3. The Niwot Ridge litter and woody debris dataset was collected from the National Ecological Observatory Network, which collectively includes 81 aquatic and terrestrial sites across 20 ecoclimatic domains. 32 of these sites sample forest litter and woody debris, and we will focus on the Niwot Ridge long-term ecological research (LTER) station in Colorado. Why might we be interested in studying litter and woody debris that falls to the ground in forests? Feel free to do a brief internet search if you feel you need more background information. 

> Answer: Litter and wood debris will drop on the ground and decompose. The nutrients of them will release to the top of soils. These procedures are important parts of the ecosystem dynamics cycle.k

4. How is litter and woody debris sampled as part of the NEON network? Read the NEON_Litterfall_UserGuide.pdf document to learn more. List three pieces of salient information about the sampling methods here: 

> Answer: The raw data like these samples' dry weight and other metadata construct the level 0 data, after quality checked, are considered Level 1 data products of NEON network.
*Individual sampling bouts construct the mass data for plant functional groups. 
*Samples are collected from elavated and group traps, respectively.
*All masses reported are reported at the spatial resolution of a single trap and the temporal resolution of a single collection event.

## Obtain basic summaries of your data (Neonics)

5. What are the dimensions of the dataset?

```{r}
dim(Neoincs)

```

6. Using the `summary` function, determine the most common effects that are studied. Why might these effects specifically be of interest?
```{r}
summary(Neoincs$Effect,5)

```

> Answer: Population and mortality are the most common effects being studied. Death is the worst outcome of a chemical and the number of dead is the most significant ecotoxicological marker.

7. Using the `summary` function, determine the six most commonly studied species in the dataset (common name). What do these species have in common, and why might they be of interest over other insects? Feel free to do a brief internet search for more information if needed.

```{r}
summary(Neoincs$Species.Common.Name,7)
```

> Answer: The six most commonly studied species are all bees in which means bees are the most vulnerable species when exposed to neonicotinoid. Most scientists studying neonicotinoid concern on the few types of bees.

8. Concentrations are always a numeric value. What is the class of Conc.1..Author. in the dataset, and why is it not numeric?

```{r}
class(Neoincs$Conc.1..Author.)
str(Neoincs$Conc.1..Author.)
```

> Answer: The type of concentration value in "Conc.1..Author" is factor. The value of it contains many characters such as: "/" and "NR" and "NR/" which not allowed in the numric type.



## Explore your data graphically (Neonics)

9. Using `geom_freqpoly`, generate a plot of the number of studies conducted by publication year. 
```{r}
ggplot(Neoincs) + geom_freqpoly(aes(x=Neoincs$Publication.Year, bins = 50))
```

10. Reproduce the same graph but now add a color aesthetic so that different Test.Location are displayed as different colors.
```{r}
ggplot(Neoincs) + geom_freqpoly(aes(x=Neoincs$Publication.Year, color =Neoincs$Test.Location), bins =50)
```

Interpret this graph. What are the most common test locations, and do they differ over time?

> Answer: From 2012 to now the most commom test location is lab, but
around 2010 the field test is more than lab. Nearly 2007, the number of test amlost is same between lab and field location. Before 2000, most test happened in field.

11. Create a bar graph of Endpoint counts. What are the two most common end points, and how are they defined? Consult the ECOTOX_CodeAppendix for more information. 

```{r}
ggplot(Neoincs) + geom_bar(aes(x=Neoincs$Endpoint)) 


```
> Answer: The two most common end points are 'NOEL' and 'LOEL'.
> NOEL, No-observable-effect-level, is defined  highest dose (concentration) producing effects not significantly  different from responses of controls according to author's reported statistical test. 
>LOEL, Lowest-observable-effect-level, is the lowest dose producing effects that were significantly different from responses of controls

## Explore your data (Litter)

12. Determine the class of collectDate. Is it a date? If not, change to a date and confirm the new class of the variable. Using the `unique` function, determine which dates litter was sampled in August 2018. 
```{r}
class(Litter$collectDate)
Litter$collectDate <- as.Date(Litter$collectDate)
class(Litter$collectDate)
unique(Litter$collectDate)
```

13. Using the `unique` function, determine how many plots were sampled at Niwot Ridge. How is the information obtained from `unique` different from that obtained from `summary`?

```{r}
unique(Litter$plotID)

```

> Answer: There are 12 plots. 'unique' provides the information about how many levels are there and no information of number, while 'summary' counts the numbers of each level.

14. Create a bar graph of functionalGroup counts. This shows you what type of litter is collected at the Niwot Ridge sites. Notice that litter types are fairly equally distributed across the Niwot Ridge sites.
```{r}
ggplot(Litter) + geom_bar(aes(Litter$functionalGroup))
```

15. Using `geom_boxplot` and `geom_violin`, create a boxplot and a violin plot of dryMass by functionalGroup. 
```{r}
ggplot(Litter) + geom_boxplot(aes(x=Litter$functionalGroup,y=Litter$dryMass))

ggplot(Litter) + geom_violin(aes(x=Litter$functionalGroup,y=Litter$dryMass),draw_quantiles = c(0.25, 0.5, 0.75))
```

Why is the boxplot a more effective visualization option than the violin plot in this case?

> Answer: The boxplot shows the rang of the data, the violin plot describes the frequency and distribution of data. When the frequency is low, the violin plot is too narrow to provide any information.  

What type(s) of litter tend to have the highest biomass at these sites? 

> Answer: Needles is the highest biomass at these sites.