---
title: "Assignment 2: Coding Basics"
author: "Xiangtian Wang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics. 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.
8. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Salk_A02_CodingBasics.Rmd") prior to submission.

The completed exercise is due on Tuesday, January 21 at 1:00 pm.


## Basics Day 1

1. Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name. 

2. Compute the mean and median of this sequence. 

3. Ask R to determine whether the mean is greater than the median. 

4. Insert comments in your code to describe what you are doing.

```{r}
#1. 
x <- seq(1,100,4)
#2. 
mean(x)
median(x)
#3. 
mean(x) > median(x)
```

## Basics Day 2

5. Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6. Label each vector with a comment on what type of vector it is.

7. Combine each of the vectors into a data frame. Assign the data frame an informative name.

8. Label the columns of your data frame with informative titles.

```{r}
s_name <- c ("Name1","Name2","Name3","Name4") # character vector
s_score <- c(90,50,40,80) # numeric vector
s_pass  <- c(TRUE, TRUE, FALSE, TRUE) # logic vector
grade <- data.frame(s_name, s_score, s_pass)
names(grade) <- c("NAME", "SCORE", "PASS");grade

```

9. QUESTION: How is this data frame different from a matrix?

> Answer: Matrice has a 2D structure that contains only on type of element. While data frame can have different modes (numric, character, logic, etc.).

10. Create a function with an if/else statement. Your function should determine whether a test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. Hint: Use `print`, not `return`. The name of your function should be informative.

11. Apply your function to the vector with test scores that you created in number 5. 

```{r}
## using ifelse
pass_test <- function(grades) {
    y <- ifelse(grades<50, "FALSE","TRUE")
    print (y)
  }
pass_test(s_score)
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer: They both worked. Using 'if' and 'else' requires more codes, even a loop for reading each value of the vector. While using "ifelse" is more convenient and concise in this problem. "ifelse" is much effective for the vector variable than "if" and "else".
