---
title: "Assignment 5: Data Visualization"
author: "Xiangtian Wang"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on Data Visualization 

## Directions
1. Change "Student Name" on line 3 (above) with your name.
3. Work through the steps, **creating code and output** that fulfill each instruction.
4. Be sure to **answer the questions** in this assignment document.
6. When you have completed the assignment, **Knit** the text and code into a single PDF file.
8. After Knitting, submit the completed exercise (PDF file) to the dropbox in Sakai. Add your last name into the file name (e.g., "Salk_A05_DataVisualization.Rmd") prior to submission.

The completed exercise is due on Tuesday, February 11 at 1:00 pm.



## Set up your session 
1. Set up your session. Verify your working directory and load the tidyverse and cowplot packages. Upload the NTL-LTER processed data files for nutrients and chemistry/physics for Peter and Paul Lakes (tidy and gathered) and the processed data file for the Niwot Ridge litter dataset. 

2. Make sure R is reading dates as date format; if not change the format to date.

```{r, message = FALSE, warning = FALSE}
#1
library(tidyverse)
library(cowplot)
library(ggplot2)
PeterPaul.chem.nutrients <- 
  read.csv("../Data/Processed/NTL-LTER_Lake_Chemistry_Nutrients_PeterPaul_Processed.csv")
PeterPaul.gather.nutrients <- 
  read.csv("../Data/Processed/NTL-LTER_Lake_Nutrients_PeterPaulGathered_Processed.csv")
Litter.mass.trap <- 
  read.csv("../Data/Processed/NEON_NIWO_Litter_mass_trap_Processed.csv")
#2
class(PeterPaul.chem.nutrients$sampledate)
class(PeterPaul.gather.nutrients$sampledate)
class(Litter.mass.trap$collectDate)
Litter.mass.trap$collectDate <- 
  as.Date(Litter.mass.trap$collectDate, format ="%Y-%m-%d")
PeterPaul.gather.nutrients$sampledate <- 
  as.Date(PeterPaul.gather.nutrients$sampledate, format = "%Y-%m-%d" )
PeterPaul.chem.nutrients$sampledate <-
  as.Date(PeterPaul.chem.nutrients$sampledate, format = "%Y-%m-%d")
```
## Define your theme
3. Build a theme and set it as your default theme.

```{r, fig.height = 3, fig.width = 4, warning = FALSE}
mytheme <- theme_classic(base_size = 12) +
  theme(axis.text = element_text(color = "black"), 
        legend.position = "top")
theme_set(mytheme)

```

## Create graphs
For numbers 4-7, create ggplot graphs and adjust aesthetics to follow best practices for data visualization. Ensure your theme, color palettes, axes, and additional aesthetics are edited accordingly.

4. [NTL-LTER] Plot total phosphorus by phosphate, with separate aesthetics for Peter and Paul lakes. Add a line of best fit and color it black. Adjust your axes to hide extreme values.

```{r}
Phosph <- ggplot (PeterPaul.chem.nutrients, aes(po4, tp_ug, color = lakename))+
              geom_point() +
              geom_smooth(method = "lm", color="black", se =F) +
              xlim(0,45) +
              xlab(expression("PO"[4]*" ("* mu*"g/ml)")) +
              ylab(expression("Tp ("*mu*"g/ml")) +
              labs(title = "Total Phosphorus by Phosphate")
print(Phosph)

```


5. [NTL-LTER] Make three separate boxplots of (a) temperature, (b) TP, and (c) TN, with month as the x axis and lake as a color aesthetic. Then, create a cowplot that combines the three graphs. Make sure that only one legend is present and that graph axes are aligned.

```{r}
box.temperature <- ggplot(PeterPaul.chem.nutrients)+
                    geom_boxplot(aes(factor(month), temperature_C, 
                                     color = lakename))
box.tp <- ggplot(PeterPaul.chem.nutrients,
                          aes(factor(month), tp_ug, color = lakename))+
                    geom_boxplot()
box.TN <- ggplot(PeterPaul.chem.nutrients,
                          aes(factor(month), tn_ug, color = lakename))+
                    geom_boxplot()
box.all <- plot_grid(box.temperature + theme(legend.position="none"),
          box.tp + theme(legend.position="none"),
          box.TN + theme(legend.position="none"),
          align = 'v', nrow = 3)
legend_box <- get_legend(box.TN + 
        theme(legend.position = "bottom"))
plot_grid(box.all, legend_box, ncol = 1, rel_heights = c(1, .2 ))

```

Question: What do you observe about the variables of interest over seasons and between lakes? 

> Answer: The temperature is highest in Summer. The nutrients are higher in Autumn than Summer, maybe due to the falling leaves. The temperature in Paul lake is a litter higher than Peter lake, while the total phosphate and total nitrogen is lower than Peter lake. 

6. [Niwot Ridge] Plot a subset of the litter dataset by displaying only the "Needles" functional group. Plot the dry mass of needle litter by date and separate by NLCD class with a color aesthetic. (no need to adjust the name of each land use)

7. [Niwot Ridge] Now, plot the same plot but with NLCD classes separated into three facets rather than separated by color. 

```{r}
#6
Litter.mass.trap.needles <- Litter.mass.trap %>%
      filter(functionalGroup=="Needles")
needles <- ggplot(Litter.mass.trap.needles)+
            geom_point(aes(collectDate,dryMass,color = nlcdClass))
print(needles)
#7
needles1 <- ggplot(Litter.mass.trap.needles)+
            geom_point(aes(collectDate,dryMass))+
            facet_grid(nlcdClass~.)
print(needles1)
```

Question: Which of these plots (6 vs. 7) do you think is more effective, and why? 

> Answer: 